---
title: "Assignment 1"
output: html_notebook
---


## Assignment 1 — Getting to know the air quality data

### Description


### Libraries and initial settings

```{r}
library(tidyverse)      # a set of packages for data transformation and visualization
library(here)
library(dplyr) # I am using this library count the number of NA values in a column as well as finding the mean and average deviation
library(ggplot2) # I am using this library to help me plot the differences between PM2.

install.packages("here")
```

```

### Exploring San Diego air quality data

#### Question 1: Counting the `NA` values

```{r}
# Count NA values in each column of aq_data_riyasha
aq_data_riyasha %>%
  summarise(across(everything(), ~ sum(is.na(.))))

```

```         
```

```         
```


    \- From the data, it shows that the number of NA missing data differs from each column with TSP having the highest missing data 2586.

#### Question 2: Statistical properties of `PM2.5`


```{r}
mean(aq_data_riyasha$PM2.5, na.rm = TRUE)
sd(aq_data_riyasha$PM2.5, na.rm = TRUE)
```

-   [**Write a brief sentence or two about your finding about the general statistics of `PM2.5`.**]{style="color: darkorange;"}

\- From the general statistics we found for PM2.5, the avaerage mean is 7.43 with a standard deviation of 3.759 which shows a fair amount of variation in the data provided.

### Visualizing Data

#### Question 3: Pollutants' relationship

-   [**Let's visualize the relationship between `PM2.5` and other pollutants. Create SIX scatter plot along with their corresponding linear trend to visualize the relationship between the other six pollutant (`TSP`, `CM`, `SD`, `NO2`, `OZ`, and `PM10`) and `PM2.5`.**]{style="color: darkorange;"}

-   [**After plotting each visualization, include a sentence or two describing the relationship you observe.**]{style="color: darkorange;"}

**PM2.5 vs TSP :** The scatter plot for PM2.5 vs TSP does not display any data point because of missing values. So, it does not show any relationship between the two variables in this plot.

**PM2.5 vs CM:** The scatter plot shows the relationship between PM2.5 and CM, the data points are scattered around the regression line. It also shows as CM increases the PM2.5 level slightly increases but the relationship is not as strong.

**PM2.5 vs SD:** The scatter plot shows a moderate psotive correlation between PM2.5 and SD. This plot shows to be stronger compared to the variable CM.

**PM2.5 vs N02:** In this scatter plot, data points are clustered towered the lower end of N02 axis with some of them increasing as NO2 increases. The relationship shows significant variability in the data.

**PM2.5 vs OZ:** In this scatter plot are densely clustered in the middle of the OZ with some spread around showing slightly positive relationship.

**PM2.5 vs PM10:** In this scatter plot it shows a strong positive relationship this can be seen by the steep slope of the regression line. As PM10 levels increases the PM2.5 also increases and shows that both are strongly associated and correlation with each other.

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = TSP, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "TSP"),
         x = "TSP",
         y = "PM2.5")

print(P)
```

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = CM, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "CM"),
         x = "CM",
         y = "PM2.5")

print(P)

```

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = SD, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "SD"),
         x = "SD",
         y = "PM2.5")

print(P)
```

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = NO2, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "NO2"),
         x = "NO2",
         y = "PM2.5")

print(P)
```

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = OZ, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "OZ"),
         x = "OZ",
         y = "PM2.5")

print(P)
```

```{r}
P <- ggplot(data = aq_data_riyasha, 
		        mapping = aes(x = PM10, 
				                  y = PM2.5))
				
P + geom_point() + geom_smooth(method = "lm")

P <- P + geom_point() + geom_smooth(method = "lm") + 
    labs(title = paste("PM2.5 vs", "PM10"),
         x = "PM10",
         y = "PM2.5")

print(P)
```
